---
title: "Movie Industry Analysis"
author: "Riley Adams, Jordan Bowman, Tracy Yu, Constanza Widel, Ruqayyah Siddique"
date: "6/6/2022"
output: html_document
---
```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Libraries
require(dplyr)
require(ggplot2)
require(GGally)
require(lmtest)
require(car)
require(leaps)
require(MASS)
require(cluster)
```

```{r, echo = FALSE}
movies_original <- read.csv("movies.csv")
dim(movies_original)

# Remove rows containing NA
movies <- movies_original[rowSums(is.na(movies_original)) == 0,]

# Create "month" using "released"
# A factor 1:12, indicating month of release
month <- numeric()

for (i in 1:length(movies$released)){
  if (grepl("January", movies$released[i])){
    month[i] <- 1
  }else if (grepl("February", movies$released[i])){
    month[i] <- 2
  }else if (grepl("March", movies$released[i])){
    month[i] <- 3
  }else if (grepl("April", movies$released[i])){
    month[i] <- 4
  }else if (grepl("May", movies$released[i])){
    month[i] <- 5
  }else if (grepl("June", movies$released[i])){
    month[i] <- 6
  }else if (grepl("July", movies$released[i])){
    month[i] <- 7
  }else if (grepl("August", movies$released[i])){
    month[i] <- 8
  }else if (grepl("September", movies$released[i])){
    month[i] <- 9
  }else if (grepl("October", movies$released[i])){
    month[i] <- 10
  }else if (grepl("November", movies$released[i])){
    month[i] <- 11
  }else{
    month[i] <- 12
  }
}

movies <- cbind(movies, month)
movies$month <- as.factor(movies$month)

# Here we can see almost 80% of the movies are from the US
# Considering IMDb is a US-based database, isolate to the US only
# sum(movies$country == "United States")/dim(movies)[1]
movies <- movies[movies$country == "United States",]

# remove variable we will not be using
drops <- c("released","director","writer","star","company")
movies <- movies[, !(names(movies) %in% drops)]
rownames(movies) <- 1:nrow(movies)
  
# Reorganize columns
organized <- c("name", "rating", "runtime", "genre", "month", "year", "votes", "score", "budget", "gross")
movies <- movies[, organized]

# Create variable for gross/budget
return_prop <-  movies$gross / movies$budget
movies <- cbind(movies, return_prop)

# All movies with blank rating were rated R
for (i in 1:nrow(movies)){
  if(movies$rating[i] == "")
    movies$rating[i] <- "R"
}

# Combine "Unrated" to "Not Rated" -> "Unrated"
for (i in 1:nrow(movies)){
  if(movies$rating[i] == "Not Rated")
    movies$rating[i] <- "Unrated"
}
```

```{r}
#check for NAs
movies_test <- movies_original[movies_original$country == "United States",]

# number of NAs
sum(rowSums(is.na(movies_test)) != 0)

# count NAs by variable
NAS_var <- colSums(is.na(movies_test))
NAS_var <- NAS_var[NAS_var != 0]
```

```{r, echo = FALSE, message = FALSE}
ggpairs(movies[,-1])
```

```{r, echo = FALSE}
hist(movies$budget, labels = TRUE,breaks = 18)
#cut at bin before 1.0e+08
mov_hist <- hist(movies$budget, labels = TRUE,breaks = 18)
budget_cat <- character()
for(i in 1:nrow(movies)){
  if (movies$budget[i] > mov_hist$breaks[5]){
    budget_cat[i] <- "high"
  }else{
    budget_cat[i] <- "low"
  }
}
movies <- cbind(movies, budget_cat)
movies$budget_cat <- as.factor(budget_cat)
#calculate percentage of movies with high budget
(159+78+65+67+44+40+9+7+9+3+2+1+1+1)/nrow(movies)
# run average release year, check overall distribution of year
mean(movies$year[movies$budget_cat == "low"])
mean(movies$year[movies$budget_cat == "high"])
# check average score between two groups (high budget/low budget)
mean(movies$score[movies$budget_cat == "low"])
mean(movies$score[movies$budget_cat == "high"])

```

Here we examine *budget* as a factor broken into two levels, based on the distribution of the histogram. Movies with a budget greater than \$80,000,000 are categorized as "high" budget while those with a budget less than or equal to \$80,000,000 are categorized as "low" budget. These two levels are stored as a factor *budget_cat*. 

The mean year of high budget movies is $2000.333$, while the mean score of high budget movies is $6.302996$. 

For low budget movies, the mean year is $2009.117$, while the mean score of low budget movies is $6.579835$.

```{r}
# ~~~~~ Responding to Q1 ~~~~~
ggpairs(movies[c("score", "votes")])
# From chart we see that the correlation is extremely significant. 
var(movies$votes)
# From above note extremely high variability
movies_q1_1 <- movies[movies$votes >= 200000,]
ggpairs(movies_q1_1[c("score", "votes")])

mean(movies$score)

mean_q1_2 <- movies[movies$votes <= 1000,]
mean(mean_q1_2$score)

mean_q1_3 <- movies[movies$votes >= 500000,]
mean(mean_q1_3$score)
```

```{r}
# ~~~~~ Responding to Q2 ~~~~~
length(unique(movies)
```

```{r}
# ~~~~~ MODEL 1 ~~~~~
# fit model and summary output
model1 <- lm(formula = score ~ rating + runtime + genre + month + year + votes + budget + gross + return_prop , data = movies)
summary(model1)
```

While some levels of the variable categorical *genre* are deemed significant predictors of *score* (e.g. *drama*, *animation*), the majority have very high p-values and are not deemed significant predictors (e.g. *thriller*, *mystery*). *Month* is not a significant predictor, nor *gross*, nor *return_prop*.


```{r}
# residual analysis
par(mfrow=c(2,2))
plot(model1)
par(mfrow=c(1,1))
```

From the Residuals vs Fitted plot, the residuals are not linear and exhibit heteroskedasticity. The QQ plot suggests that the residuals are not normally distributed as the lower and higher standardized residuals do not line up with the theoretical quantiles. We do have some outliers as well as one particularly high leverage point observations 2817 (The 2007 movie "Paranormal Activity").

```{r}
# Model Selection

## Forward selection 
model1_forward <- regsubsets(score ~ rating + runtime + genre + month + year + votes + budget + gross + return_prop, 
                           data = movies, method = "forward")
cbind(summary(model1_forward)$which, "adjusted r^2" = summary(model1_forward)$adjr2)

## Backward selection 
model1_backward <- regsubsets(score ~ rating + runtime + genre + month + year + votes + budget + gross + return_prop, 
                           data = movies, method = "backward")
cbind(summary(model1_backward)$which, "adjusted r^2" = summary(model1_backward)$adjr2)
```

Using a step-wise model selection procedure with $R^2_{adj}$ as a selection criterion, we will choose a model which produces the maximum value for $R^2_{adj}$. This method will implement partial F-tests and then one-by-one eliminate the least significant predictor. In this case, per both backward and forward elimination, the suggested model is score ~ ratingR, runtime, genreAnimation, genreBiography, genreDrama, genreHorror, votes, budget. 

Since only 1 of 8 levels for *rating* is deemed significant, and only 4 of 15 levels for *genre* is deemed significant. We will drop these predictors along with *return_prop*, *gross*, *year*, and *month* from the model and analyze the model again.

```{r}
# ~~~~~ MODEL 2 ~~~~~~~~~
# fit model and summary output
model2 <- lm(formula = score ~ runtime + votes + budget, data = movies)
summary(model2)
```

```{r}
# residual analysis
par(mfrow=c(2,2))
plot(model2)
par(mfrow=c(1,1))
```

--[[EXPLAIN BELOW]]--
Diagnosis:

Residuals vs Fitted: non-linearity, heteroskedasticity

QQ-Plot: non-normality

scale-location: possible outliers 

Residuals vs Leverage: high leverage points (1148, 2889,1147)

Remedy:

Will begin by attempting boxcox transformation to address nonlinearity and nonconstant variance in residuals.

```{r}
# ~~~~ BoxCox Transformation ~~~~~~
# Boxcox for model2
boxcox(model2)
```

Since the boxcox plot has its maximum value near $\lambda = 2$, we raise $\hat{score}$ to the 2nd power. We fit the new model (model3) as such:

$\hat{score} = \sqrt{\hat\beta_0 + \hat\beta_{\text{runtime}} \cdot \text{runtime} + \hat\beta_{\text{votes}} \cdot \text{votes} + \hat\beta_{\text{budget}} \cdot \text{budget}}$

```{r}
# ~~~~~ MODEL 3 ~~~~~~~~~
# square outcome variable (score)
score_sq <- (movies$score)^2
# fit model and summary output
model3 <- lm(formula = score_sq ~ runtime + votes + budget, data = movies)
summary(model3)
```

All predictors from model2 are still significant. Adjusted R-squared has increase by $0.3945 - 0.3372 = 0.0573$.

```{r}
# residual analysis
plot(model3)
```

The transformation has only marginally improved the residual plots. The same issues remain. We can try removing the high leverage points from the model. (Observations 1148,2889,1147).

```{r}
# ~~~~~ MODEL 4 ~~~~~~~~~

# create vector containing highest leverage points.
# (Pulp Fiction, The Dark Knight, The Shawshank Redemption)
highlev <- c(1148,2889,1147)
movies[highlev,]

# new dataframe without high leverage points
movies_lowlev <- movies[-highlev,]
rownames(movies_lowlev) <- 1:nrow(movies_lowlev)

# have to remove same values in score_sq
score_sq_ll <- score_sq[-highlev]

# fit model and summary output
model4 <- lm(formula = score_sq_ll ~ runtime + votes + budget, data = movies_lowlev)
summary(model4)
```

All predictors still significant. Adjusted R-squared improved by $0.3968 - 0.3945 = 0.0023$.

```{r}
# residual analysis
plot(model4)
```

```{r}
# high leverage
hl_2 <- c(1741,1147,3149)
movies_lowlev[hl_2,]
# seems like the movies that keep throwing off the model are those that are "big hits" or "classics"
```

Seems like the movies that keep throwing off the model are those that are "big hits" or "classics". Will, again, remove these observations.

```{r}
# ~~~~~ MODEL 5 ~~~~~~~~~

# new dataframe without high leverage points
movies_lowlev2 <- movies_lowlev[-hl_2,]
rownames(movies_lowlev2) <- 1:nrow(movies_lowlev2)

# have to remove same values in score_sq_ll
score_sq_ll2 <- score_sq_ll[-hl_2]

# fit model and summary output
model5 <- lm(formula = score_sq_ll2 ~ runtime + votes + budget, data = movies_lowlev2)
summary(model5)
```

All predictors still significant. Adjusted R-squared improved by $0.3984 - 0.3968 = 0.0016$.

```{r}
# residual analysis
plot(model5)
```

```{r}
# high leverage
hl_3 <- c(3409,3668,1256)
movies_lowlev2[hl_3,]

boxcox(model5)
```

More of those well-loved, classic movies throwing off the model. Maybe movie magic does exist. Our model may work for those movies that don't have the "magic". Residuals are getting better. Want to try transforming one more time, box cox suggests raising to the second power again.

```{r}
# ~~~~~ MODEL 6 ~~~~~~~~~
score_4th <- (score_sq_ll2)^2

# fit model and summary output
model6 <- lm(formula = score_4th ~ runtime + votes + budget, data = movies_lowlev2)
summary(model6)
```

Adjusted R-squared increased by $0.4844 - 0.3984 = 0.086$

```{r}
# residual analysis
plot(model6)
```

Residuals are looking way better. Linearity improved, not quite there but *much* closer. Homoskedasticity appears more likely. Residual Quantiles are quite close to theoretical quantiles, could pass for normality.

```{r}
# ~~~~~~~~ Residual Analysis: Formal Tests ~~~~~~~~

# Shapiro-Wilk Test
# test for normality
# H0: data are normally distributed
shapiro.test(resid(model6))
# we reject the null, so the residuals are not normal

# Box-Pierce Test
# test for correlated errors
# H0: The residuals are independently distributed
Box.test(model6$residuals)
# We reject the null, the residuals are independently distributed

# Breusch-Pagan Test
# test for homoskedasticity
# H0: the errors are homoskedastic
bptest(model6)
# We reject the null, the residuals are heteroskedastic (non constant variance)

```

```{r}
# Gower dist: https://arxiv.org/ftp/arxiv/papers/2101/2101.02481.pdf
movies_clust <- movies[-1]
movies_clust$rating <- as.factor(movies_clust$rating)
movies_clust$genre <- as.factor(movies_clust$genre)
movies_clust$month <- as.factor(movies_clust$month)

gow_dist_full <- daisy(movies_clust, metric = "gower")
gow_full_hclust <- hclust(gow_dist_full)
plot(gow_full_hclust)
cut_gow_full_hclust <- cutree(gow_full_hclust, k = 3)
movies[cut_gow_full_hclust == 3, 1]

gow_dist_intuitive <- daisy(movies_clust[c(1, 3, 5, 7, 8)], metric = "gower")
gow_intuitive_hclust <- hclust(gow_dist_intuitive)
plot(gow_intuitive_hclust)
abline(h = 0.72, col = "red")
cut_gow_int_hclust <- cutree(gow_int_hclust, k = 6)
movies[cut_gow_int_hclust == 6, 1]
```




















